For the back-end side of my application, I chose a tech stack that I thoroughly enjoy working with: Laravel, MariaDB, and Filament, all fully dockerized and running on an Ubuntu VM hosted on my personal home server.

\begin{itemize}
    \item Laravel: As a PHP framework, Laravel offers a robust and elegant solution for building scalable, secure, and maintainable back-end applications. It provides built-in features like routing, ORM (Eloquent), middleware, and more, allowing me to focus on the business logic of Mosaico rather than boilerplate code.
    \item MariaDB: For the database, I selected MariaDB because it is a community-developed fork of MySQL and provides excellent performance, scalability, and open-source benefits. Its compatibility with MySQL made it a straightforward choice for integration with Laravel's Eloquent ORM.
    \item Filament: As an admin panel generator, Filament enables me to create highly interactive admin interfaces with minimal effort. It integrates seamlessly with Laravel, allowing me to manage widgets and user data from a sleek UI while offering powerful tools like forms, tables, and resource management.
    \item Dockerization: To ensure the application is easy to maintain and deploy, I containerized the entire stack using Docker. Dockerization provides isolation for each component and eliminates the "it works on my machine" problem by creating consistent environments across development and production.
    \item Ubuntu VM: The server runs on an Ubuntu virtual machine hosted on my personal home server. This setup gives me full control over the deployment process and allows for flexibility in configuring and maintaining the environment.
\end{itemize}

This stack not only aligns with my technical preferences but also ensures that Mosaico's back-end is scalable, flexible, and easy to manage, making it a solid foundation for the future development of the platform.

\section{Widget manager}
\input{tesi/chapters/8/widget-manager}

\section{API}
\input{tesi/chapters/8/api}
\newpage
\section{Landing Page}
\input{tesi/chapters/8/landing-page}
\newpage
\section{Documentation}
\input{tesi/chapters/8/documentation}